---
title: "isdbayes_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{isdbayes_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(isdbayes)
library(here)
source(here("R/paretocounts.R"))
library(brms)
library(tidyverse)
```

# simulate a single size distribution
This simulates data with rparetocounts(), which takes values of n, mu, vreal2, and vreal3. n is the number of body sizes to simulate, mu is lambda, vreal2 is xmin, and vreal3 is xmax. The use of "mu" and "vreal2" and "vreal3" is required by brms to represent the quantities of interest, so those terms are used in this package (which can be confusing).

```{r}
isd_data = tibble(x = rparetocounts()) %>% 
  mutate(xmin = min(x), # xmin
         xmax = max(x)) %>% #xmax
  group_by(x) %>% 
  add_tally(name = "counts") # get counts of each body size. If all sizes are unique, this is just 1. This is where size densities would also go
```

# estimate lambda
```{r}
fit = brm(x | vreal(counts, xmin, xmax) ~ 1, 
          data = isd_data,
          stanvars = stanvars,
          family = paretocounts(),
          chains = 1, iter = 1000)
```

# simulate multiple size distributions
```{r}

x1 = rparetocounts(mu = -1.8) # `mu` is required wording from brms. in this case it really means the lambda exponent of the ISD
x2 = rparetocounts(mu = -1.5)
x3 = rparetocounts(mu = -1.2)

isd_data = tibble(x1 = x1,
                  x2 = x2,
                  x3 = x3) %>% 
  pivot_longer(cols = everything(), names_to = "group", values_to = "x") %>% 
  group_by(group) %>% 
  mutate(xmin = min(x),
         xmax = max(x)) %>% 
  group_by(group, x) %>% 
  add_count(name = "counts")
```

# fit multiple size distributions with a fixed factor
```{r}
fit2 = brm(x | vreal(counts, xmin, xmax) ~ group, 
           data = isd_data,
           stanvars = stanvars,
           family = paretocounts(),
           chains = 1, iter = 1000)
```

# fit multiple size distributions with a varying intercept
```{r}
fit2 = brm(x | vreal(counts, xmin, xmax) ~ (1|group), 
           data = isd_data,
           stanvars = stanvars,
           family = paretocounts(),
           chains = 1, iter = 1000)
```



